noam_data_exploration_v1

### 2024-08-16 noam_data_exploration_v3.Rmd
re-ran models at island side level. Began figuring out bray curtis/beta diversity stuff at quad level. Still WIP.

(richness_quad_plot <-
  ggplot(data = alpha_diversity_quad_means) +
  geom_point(aes(x = year, y = richness_mean, colour = site, shape = habitat, group = site_habitat), size = 8) +
  geom_line(aes(x = year, y = richness_mean, colour = site, group = site_habitat), size = 1.2) +
  geom_errorbar(aes(x = year, ymin = richness_mean - richness_se , ymax = richness_mean + richness_se, colour = site,
                    group = site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat)+
  labs(title = "a. richness quad level\n", y = "richness\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

@@

columns <- c("location", "Intercept", "Slope") # names of columns for dataframe
gls_coefficients <- # new dataframe name
  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$location)), # 1200 quadrats = 1200 different time series
                    ncol = length(columns)) # specified to make sure number of columns doesn't exceed what was specified
             )
colnames(gls_coefficients) <- columns # change column names to what was specified
gls_coefficients$location <- levels(alpha_diversity_quad$location) # populate location column with each quad/time series

@@

  # populate appropriate row and column with the relevant coeffcients from gls
  gls_coefficients[i, 2] <- unname(coef(loop_mod)[1])
  gls_coefficients[i, 3] <- unname(coef(loop_mod)[2])

@@

hist(gls_coefficients$Intercept) # same with intercept - although it probably shouldn't be

@@

slope_emm <- emmeans(slope_mod, ~ habitat*site) # create emmeans object
emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
# Note: not sure I want to include this
emmeans(slope_mod, pairwise ~ site | habitat)
```

```{r gls plot, fig.height=20, fig.width=20}
# not sure how to visualize here... thinking something like Dornelas et al. 2014 Fig 2a
# I reached out to them about getting their code for this

ggplot(data = gls_coefficients_merged) +
  geom_histogram(aes(x = Slope), bins = 30) +
  geom_vline(xintercept = 0, size = 2, linetype = 2) +
  facet_grid(habitat ~ site) +
#  theme_classic() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain")
  )

@@

```{r gls intercept, fig.height=20, fig.width=20}
intercept_mod <- glmmTMB(Intercept ~ habitat*island_side + (1|site/transect), data = gls_coefficients_merged)
summary(intercept_mod) # a lot of variance w/i transect, less so w/i site
plot(gls_int_res <- simulateResiduals(intercept_mod)) # same issues as above

hist(residuals(intercept_mod))
plot(predict(intercept_mod) ~ residuals(intercept_mod))
Anova(intercept_mod)
#                        Chisq Df Pr(>Chisq)    
# habitat             1307.838  3  < 2.2e-16 ***
# island_side           26.358  2  1.889e-06 ***
# habitat:island_side  789.627  6  < 2.2e-16 ***
r.squaredGLMM(intercept_mod) # 0.16, 0.18
plot(allEffects(intercept_mod))

ggplot(data = gls_coefficients_merged) +
  geom_histogram(aes(x = Intercept), bins = 30) +
  geom_vline(xintercept = 0, size = 2, linetype = 2) +
  facet_grid(habitat ~ island_side) +
#  theme_classic() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain")
  )

@@

# probably the closest to what others have done:
richness_mod_2 <- glmmTMB(
  richness ~ year_factor*habitat*site + (1 | site/transect/location) + ar1(year_factor + 0 | location),
  family = poisson,
  data = alpha_diversity_quad,
  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)), na.action = na.pass
) # overspecified but going with it for now

@@

top_rich_mod <- glmmTMB(richness ~ year_factor*habitat + year_factor*site + (1 | site/transect/location) + ar1(year_factor + 0 | location),   family = poisson, data = alpha_diversity_quad, control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)))
summary(top_rich_mod)
Anova(top_rich_mod)
```
```{r turnover troubleshooting, fig.height = 30, fig.width=25}

@@

# model
total_turnover_mod1 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site +
                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
                               family = beta_family(), data = total_turnover)
summary(total_turnover_mod1) # low temporal autocorrelation... maybe don't need ar1 term?
plot(total_turnover_res <- simulateResiduals(total_turnover_mod1)) # residuals look very weird
Anova(total_turnover_mod1) # obviously everything sig

total_turnover_mod2 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site + (1 | site/transect/location),
                               family = beta_family(), data = total_turnover)
summary(total_turnover_mod2) # 
plot(total_turnover_res2 <- simulateResiduals(total_turnover_mod2)) # residuals still look like shit
Anova(total_turnover_mod2) 
anova(total_turnover_mod2, total_turnover_mod1) # models are sig different from each other
AICc(total_turnover_mod2, total_turnover_mod1) # first model is better
# need to figure out residuals problem
plot(allEffects(total_turnover_mod1))

@@

(turnover_plot <-
  ggplot(data = total_turnover_means) +
  geom_point(aes(x = year, y = Turnover_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = year, y = Turnover_mean, colour = site, group = site_habitat), size = 1.2) +
  geom_errorbar(aes(x = year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = site,
                    group = site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site_habitat, nrow = 6)+
  labs(title = "a. Turnover\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
#
```

@@

```{r nmds, fig.height = 15, fig.width = 15}
species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
species_matrix <- species_matrix/100 # convert from percent to proportion
rownames(species_matrix) <- paste(data$year, data$site_habitat, data$transect)
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )

# identifying species w/ > 10% cover for plotting later
# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 0.1))
abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
abundant_species_names <- colnames(abundant_species_matrix)

>>>>

(richness_quad_plot <-
  ggplot(data = alpha_diversity_quad_means) +
  geom_point(aes(x = year, y = richness_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8) +
  geom_line(aes(x = year, y = richness_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
  geom_errorbar(aes(x = year, ymin = richness_mean - richness_se , ymax = richness_mean + richness_se, colour = island_side,
                    group = island_side_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = island_side_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat)+
  labs(title = "a. richness quad level\n", y = "richness\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

@@

columns <- c("location", "Intercept", "Slope", "Phi") # names of columns for dataframe
gls_coefficients <- # new dataframe name
  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$location)), # 1200 quadrats = 1200 different time series
                    ncol = length(columns)) # specified to make sure number of columns doesn't exceed what was specified
             )
colnames(gls_coefficients) <- columns # change column names to what was specified
gls_coefficients$location <- levels(alpha_diversity_quad$location) # populate location column with each quad/time series

@@

  # populate appropriate row and column with the relevant coeffcients from gls
  gls_coefficients[i, 2] <- unname(coef(loop_mod)[1])
  gls_coefficients[i, 3] <- unname(coef(loop_mod)[2])
  gls_coefficients[i, 4] <- loop_mod[["modelStruct"]][["corStruct"]]

@@

hist(gls_coefficients$Intercept) # same with intercept - although it probably shouldn't be
hist(gls_coefficients$Phi)

@@

slope_emm <- emmeans(slope_mod, ~ habitat*site) # create emmeans object
emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
# Note: not sure I want to include this
emmeans(slope_mod, pairwise ~ site | habitat)

## AT COAST-LEVEL
slope_mod <- glmmTMB(Slope ~ habitat*island_side + (1|island_side/site/transect), data = gls_coefficients_merged)
summary(slope_mod) # pretty low variance in random effects
plot(gls_res <- simulateResiduals(slope_mod))

hist(residuals(slope_mod))
plot(predict(slope_mod) ~ residuals(slope_mod))
Anova(slope_mod)
#                        Chisq Df Pr(>Chisq)    
# habitat             1301.246  3  < 2.2e-16 ***
# island_side           16.075  2  0.0003231 ***
# habitat:island_side  799.999  6  < 2.2e-16 ***
r.squaredGLMM(slope_mod) # 0.16, 0.18
plot(allEffects(slope_mod))
# backreef: richness increases on east and north shore, but relatively constant west side
# Fringe: richness decreasing on east and west side, but relatvely constant on north shore
# outer_10: richness neutral/decreasing on east side, but neutral/increasing on north and west side
# outer_17: richness decreasing north/east side, but neutral west side
# Worth noting trends are slight - arguable whether they are biologically meaningful

slope_emm <- emmeans(slope_mod, ~ habitat*island_side) # create emmeans object
emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
# Note: not sure I want to include this
emmeans(slope_mod, pairwise ~ island_side | habitat)
# north and east coast not different from each other in backreef
# east and west not different from each other in fringing
# north and west not different from each other outer 10
# no sites different from each other outer 17
slope_est <- as.data.frame(slope_emm) # create df of emmeans estimates

@@

```{r gls plot, fig.height=20, fig.width=25}
# not sure how to visualize here... thinking something like Dornelas et al. 2014 Fig 2a
# I reached out to them about getting their code for this

ggplot(data = gls_coefficients_merged, aes(x = Slope)) +
  geom_histogram(bins = 40, aes(y = after_stat(ndensity))) +
  geom_density(alpha=0,aes(y = after_stat(ndensity)), size = 1.5 ) +
  geom_vline(xintercept = 0, size = 2, linetype = 2) +
  geom_ribbon(data = merge(gls_coefficients_merged, slope_est), aes(xmin = lower.CL, xmax = upper.CL, ymin = 0, ymax = Inf),
              fill = "red", alpha = 0.1)+
  geom_vline(data = slope_est, aes(xintercept = emmean), size = 2, linetype = 1, colour = "red") +
  facet_grid(island_side ~ habitat) +
  theme_classic() +
  labs(y = "Density\n") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain")
  )

```{r gls intercept, fig.height=20, fig.width=25}
intercept_mod <- glmmTMB(Intercept ~ habitat*island_side + (1|island_side/site/transect), data = gls_coefficients_merged)
summary(intercept_mod) # a lot of variance w/i transect, less so w/i site
plot(gls_int_res <- simulateResiduals(intercept_mod)) # same issues as above

hist(residuals(intercept_mod))
plot(predict(intercept_mod) ~ residuals(intercept_mod))
Anova(intercept_mod)
#                        Chisq Df Pr(>Chisq)    
# habitat             1285.762  3  < 2.2e-16 ***
# island_side           15.644  2  0.0004009 ***
# habitat:island_side  776.312  6  < 2.2e-16 ***
r.squaredGLMM(intercept_mod) # 0.16, 0.17
plot(allEffects(intercept_mod))

int_emm <- emmeans(intercept_mod, ~ habitat*island_side) # create emmeans object
#emmeans::contrast(int_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")

emmeans(intercept_mod, pairwise ~ island_side | habitat)
# same trends as with slope

int_est <- as.data.frame(int_emm) # create df of emmeans estimates

ggplot(data = gls_coefficients_merged, aes(x = Intercept)) +
  geom_histogram(bins = 40, aes(y = after_stat(ndensity))) +
  geom_density(alpha=0,aes(y = after_stat(ndensity)), size = 1.5 ) +
  geom_vline(xintercept = 0, size = 2, linetype = 2) +
  geom_ribbon(data = merge(gls_coefficients_merged, int_est), aes(xmin = lower.CL, xmax = upper.CL, ymin = 0, ymax = Inf),
              fill = "red", alpha = 0.1)+
  geom_vline(data = int_est, aes(xintercept = emmean), size = 2, linetype = 1, colour = "red") +
  facet_grid(island_side ~ habitat) +
  theme_classic() +
  labs(y = "Density\n") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain")
  )

@@

# probably the closest to what others have done:
richness_mod_2 <- glmmTMB(
  richness ~ year_factor*habitat*island_side + (1 | island_side/transect/location) + ar1(year_factor + 0 | location),
  family = poisson,
  data = alpha_diversity_quad,
  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)), na.action = na.pass
) # overspecified but going with it for now

@@

# top_rich_mod <- glmmTMB(richness ~ year_factor*habitat + year_factor*island_side + (1 | island_side/transect/location) + # ar1(year_factor + 0 | location),   family = poisson, data = alpha_diversity_quad, control = glmmTMBControl(optCtrl = list(iter.max = # 2000, eval.max = 2000)))
# summary(top_rich_mod)
# Anova(top_rich_mod)

# Julianna's model structure
j_rich_mod <- glmmTMB(richness ~ year + habitat + (1|site/transect),
               family = poisson,
               data = alpha_diversity_quad)
summary(j_rich_mod)
plot(allEffects(j_rich_mod))

grid <- expand_grid(year = unique(alpha_diversity_quad$year),
                       habitat = unique(alpha_diversity_quad$habitat),
                       site = unique(alpha_diversity_quad$site),
                       island_side = unique(alpha_diversity_quad$island_side),
                       transect = unique(alpha_diversity_quad$transect))
grid$Preds <- predict(j_rich_mod, newdata = grid)

grid %>% 
  # get transect means
  group_by(year, habitat, island_side) %>% 
  summarize(Pred_mean_richness = mean(Preds),
            SD_richness = sd(Preds)) %>% 
  ggplot(aes(x = year, y = Pred_mean_richness, group = habitat, color = habitat)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~island_side)

# Given that this is largely looking at slopes for temporal trends in adiv, I prefer the gls
```

```{r turnover troubleshooting, fig.height = 30, fig.width=25}

@@

# model
total_turnover_mod1 <- glmmTMB(Turnover_trans ~ year_factor*habitat*island_side +
                                 (1 | island_side/site/transect) + ar1(year_factor + 0 | location),
                               family = beta_family(), data = total_turnover)
summary(total_turnover_mod1) # low temporal autocorrelation... maybe don't need ar1 term?
plot(total_turnover_res <- simulateResiduals(total_turnover_mod1)) # residuals look very weird
Anova(total_turnover_mod1) # obviously everything sig

# total_turnover_mod2 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site + (1 | site/transect/location),
#                                family = beta_family(), data = total_turnover)
# summary(total_turnover_mod2) # 
# plot(total_turnover_res2 <- simulateResiduals(total_turnover_mod2)) # residuals still look like shit
# Anova(total_turnover_mod2) 
# anova(total_turnover_mod2, total_turnover_mod1) # models are sig different from each other
# AICc(total_turnover_mod2, total_turnover_mod1) # first model is better
# # need to figure out residuals problem
# plot(allEffects(total_turnover_mod1))
#

@@

(turnover_plot <-
  ggplot(data = total_turnover_means) +
  geom_point(aes(x = year, y = Turnover_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = year, y = Turnover_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
  geom_errorbar(aes(x = year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = island_side,
                    group = island_side_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = island_side_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~island_side_habitat, nrow = 6)+
  labs(title = "a. Turnover\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
#
```

@@

```{r bray curtis set up}
species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
species_matrix <- species_matrix/100 # convert from percent to proportion
rownames(species_matrix) <- data$location
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
# data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
brayfull <- vegdist(species_matrix, "bray")
# View(brayfull)

# playing around with stuff here - ignore for now
x <- subset(data_subset, location == "lter_1_backreef_algae_transect_1_quad_1")
year_combinations <- expand.grid(Year1 = x$year, Year2 = x$year)
year_combinations <- year_combinations[year_combinations$Year1 != year_combinations$Year2, ] # filter out year 1 == year 2
year_combinations$Year_Difference <- abs(year_combinations$Year1 - year_combinations$Year2) # find difference in years
year_combinations$location <- rep("lter_1_backreef_algae_transect_1_quad_1", nrow(year_combinations))
year_combinations$bray <- NA
for(i in 1:nrow(year_combinations)){
  subset(x)
}

```{r old nmds code, fig.height = 15, fig.width = 15, eval = FALSE}
species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
species_matrix <- species_matrix/100 # convert from percent to proportion
rownames(species_matrix) <- paste(data$year, data$site_habitat, data$transect)
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
# identifying species w/ > 10% cover for plotting later
# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 0.1))
abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
abundant_species_names <- colnames(abundant_species_matrix)
